Shaping Anti-Corruption Campaigns in Nigeria

### 2024-12-03 paper/paper.qmd
prose checks

---
title: "Perceptions of Corruption in Nigeria"
subtitle: "Rural Nigerians and Less Educated  Groups Report Higher Corruption"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-"
date: today
date-format: long
abstract: "Peceptions of corruption in the Nigerian presidency vary significantly across demogrpahic groups. Using Afrobarometer survey data and Bayesisan ordinal logistic regression, this study looks at how education, gender, age, and urban versus rural residence influence these perceptions. The results show that urban residents and yourger individuals are more likely to perceive high corruption, while rura populations and those with higher education show views that are moderate. These results show the need for targeted anti-corruption campaigns to adress demographic differences and imporve governance accountability in Nigeria"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

#install.packages("modelsummary")
#install.packages("rstanarm")
#install.packages("dplyr")
#install.packages("here")
#install.packages("arrow")
#install.packages("tinytable")
library(modelsummary)
library(rstanarm)
library(dplyr)
library(here)
library(arrow)
library(bayesplot)
library(ggplot2)
library(arrow)
library(here)
library(patchwork)

@@

Understanding these demographic differences is important for designing effective anti-corruption campaigns. Making sure outreach efforts are tailored to the unique experiences and perceptions of different demographic groups can increase their effectiveness and increase public trust. For example, youth-focused campaigns that leverage social media platforms and urban outreach using investigative journalism can engage the people. At the same time, localized, culturally appropriate campaigns in rural areas can address the different needs of these populations. A the UNODC emphasizes, 'targeted and effective anti-corruption policies will improve likes of Nigerian citizens" [@afrobarometer2023].

The remainder of this paper is structured as follows; @sec-data detailed the data used for this paper, @sec-model introduces the Bayesian ordinal logistic regression model used to study the relationship between our variables of interest. @sec-results presents the findings from the model, detailing how different demographic factors like education and age affect corruption perception in Nigeria. @sec-discussion examines these results, discussing their implications for anti-corruption policies, weaknesses and next steps. @sec-survey-appendix explores the details of the Afrobarometer survey. @sec-model-details shows plots for Posterior predictive checks and Diagnostics as well as their analysis. @sec-survey-appendix details the survey methodology used in the Afrobaromenter data and @sec-ethics addresses key ethical principles and the ethical challenges of surveying sensitive topics like corruption.

@@

Our primary measurement task is to capture public opinion towards
economic, political and socioeconomic issues and translate it into
actionable lintel. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses.

@@

Each survey question acts as a data proxy for these complex real-world
dynamics, reducing them to quantifiable variables while maintaining
their conceptual integrity.

This dataset captures a unique moment in Nigeria's socio-political
landscape, reflecting how citizens perceive democracy and governance in
the context of ongoing economic and political challenges. Alternative
datasets, such as Nigeria’s national statistics or World Bank survey
data, could provide economic indicators but lack the granularity of
Afrobarometer’s public opinion data.

The data for this study was downloaded, cleaned, analyzed, modeled, and
visualized using R [@citeR], a statistical programming language.[@afrobarometer2023] served as the primary data source, providing valuable survey data for analysis.***Telling Stories with Data*** [@tellingstories] was referenced for its code and methodologies. The
following packages were used:

@@

#install.packages("knitr")
#install.packages("kableExtra")
library(knitr)
library(kableExtra)

@@

The cleaned dataset used in this study is summarised in
@tbl-summary-statistics, which provides key statistics to offer an
overview of the dataset's structure and characteristics. The most
frequent age group among respondents is identified, representing the
predominant demographic in terms of age. The table also highlights the
urban and male percentages, which indicate the proportion of urban
residents (43.38%) and male respondents (49.68%) in the sample.

Further, the table breaks down the perceptions of corruption into four
categories—"None," "Some," "High," and "All"—with counts for each,
offering insights into respondents' views on corruption in the
presidency. Additionally, the table details the distribution of
respondents across education levels, including categories such as
"Primary," "Secondary," "University," and "Postgraduate." Together,
these summary statistics provide a comprehensive snapshot of the
dataset, enabling a clear understanding of its demographic and
categorical variables.

@@

The heatmap reveals that respondents who completed secondary school or
some secondary schooling dominate the High Corruption and Some
Corruption categories. Respondents with higher education levels
(university and post-graduate) are distributed more evenly across the
corruption perception categories, with fewer concentrated in any
specific category. Conversely, respondents with no formal schooling or
only informal education are sparsely represented across all
categories.There is a notable concentration of corruption perceptions
among respondents with mid-level education (secondary schooling), while
those with higher education tend to have more varied perceptions. These
findings suggest a potential link between educational attainment and
experiences or attitudes toward corruption, particularly among those
with secondary-level education.

@@

The chart reveals that rural and urban respondents share similar
patterns of corruption perception. The majority of respondents in both
groups fall under the Some Corruption and High Corruption categories,
with Some Corruption being slightly more prevalent. The All Corrupt
category is the third most common perception in both groups, while the
No Corruption category is consistently the least represented.

@@

#install.packages("broom.mixed")
#install.packages("kableExtra")

@@

Higher education levels consistently show positive probabilities,
indicating that individuals with more advanced education levels are more
likely to perceive higher corruption levels. Living in rural areas has a
significantly negative probability, suggesting that rural residents are
less likely to perceive high corruption compared to urban residents.
Gender shows a moderate positive probability, with females slightly more
likely to perceive higher corruption levels than males. There seems to
be minimal variation observed across age groups, suggesting that age
does not substantially influence perceptions of corruption.

@@

Urban respondents are somewhat more likely than their rural counterparts
to perceive high corruption in the presidency. This is consistent with
findings that urban populations have greater access to investigative
reporting and public discourse on corruption, which can amplify
awareness and perceptions [@u4nigeria2023]. On the other hand, rural
respondents, who may rely on localized patronage networks for essential
services, often exhibit more neutral or unclear perceptions of
corruption, influenced by limited media access and dependency on these
informal systems. Campaigns should focus on leveraging urban platforms
like social media, community-based programs, and public dialogues to
engage critical urban audiences, while also expanding rural outreach
through local radio, grassroots organizations, and educational
interventions tailored to these settings.

@@

This paper is focused on looking at trends in the corruption perception
between different demographics and seeing how these differences can be
leveraged to maximize anti corruption campaigns and policy in
Nigeria.From this paper, we find different actionable strategies for
improving the effectiveness of anti-corruption campaigns in Nigeria.
First, targeting rural and urban populations differently might be
effective. Urban residents, who have greater access to investigative
media and public debates, are more likely to perceive high corruption.
Campaigns in urban areas should therefore focus on digital platforms
like social media and mobile apps, as well as partnerships with
investigative journalists and public accountability platforms to engage
this group effectively. In rural areas, where reliance on patronage
systems and limited media access shape perceptions, grassroots
communication strategies like community radio, local town halls, and
collaborations with trusted community leaders are better suited to
disseminate anti-corruption messaging.

Education also emerges as a key determinant of corruption perceptions.
Secondary school graduates exhibit the highest levels of awareness, this
suggests that civic education campaigns should focus on this demographic
to harness their increased engagement. Integrating anti-corruption
education into school curricula could early awareness and long-term
attitudinal changes.Youth engagement is very important, given that
younger age groups, particularly those aged 25–34, report the highest
levels of dissatisfaction with corruption. Social media campaigns
tailored to platforms popular with youth, together with interactive town
hall discussions and youth-led grassroots initiatives, could channel
this demographic's energy into organized anti-corruption efforts.

Lastly, rebuilding public trust in institutions should be a priority.
Transparent hiring practices, visible enforcement of anti-corruption
laws, and strengthened whistleblower protections are important steps to
showing institutional accountability. Localized outreach through
community leaders can also help bridge the trust gap, particularly in
rural settings. Policymakers should put these findings into a
data-driven framework so they can adapt strategies based on demographic
trends and feedback.

@@

- Africa-Wide Generalization: While Afrobarometer provides national-level insights, its findings cannot be generalized to the entire continent due to contextual variations.

@@

- **Inclusion of Marginalized Populations**: Future iterations could include anonymous digital surveys to capture insights from institutionalized or transient populations.
- **Addressing Non-Response Bias**: Enhanced recruitment strategies, such as partnerships with community organizations, could increase participation among underrepresented groups.
- **Community Engagement**: Sharing results directly with communities through accessible formats (e.g., infographics or local town halls) would increase transparency and accountability.

@@

The ethical considerations underpinning Afrobarometer’s methodology ensure the credibility and validity of its findings while safeguarding respondent dignity. However, addressing remaining gaps—such as the exclusion of certain populations and challenges in conflict zones—would further enhance the ethical rigor of corruption perception research in Africa. By adhering to established ethical frameworks and remaining responsive to emerging challenges, studies like this can continue to provide valuable insights while maintaining the highest ethical standards.To do my part in ensuring ethical standards are upheld, i have included a datasheet for the Afrobarometer dataset.

@@

```{r}
#| label: fig-variables-of-interest-heatmap
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "combined variables of interest"

#install.packages("tidyr")
library(tidyr)
# Combine data for heatmap
heatmap_data <- cleaned_data %>%
  pivot_longer(
    cols = c(URBRUR, corruption_presidency, education, age_group),
    names_to = "variable",
    values_to = "category"
  ) %>%
  count(variable, category) %>%
  mutate(variable = factor(variable, levels = c("URBRUR", "corruption_presidency", "education", "age_group")))

# Create heatmap
ggplot(heatmap_data, aes(x = category, y = variable, fill = n)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(
    title = "Heatmap of Variable Distributions",
    x = "Category",
    y = "Variable",
    fill = "Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

![Combined plot](combined_plot.png){width="120%"}

>>>>

---
title: "Shaping Anti-Corruption Campaigns in Nigeria"
subtitle: "Youth Populations, Urban Areas and Moderately Educated Citizens Are Important for Anti-Corruption Campaigns "
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-"
date: today
date-format: long
abstract: "Peceptions of corruption in the Nigerian presidency vary significantly across demogrpahic groups. Using Afrobarometer survey data and Bayesisan ordinal logistic regression, this study looks at how education, gender, age, and urban versus rural residence influence these perceptions. The results show that urban residents and yourger individuals are more likely to perceive high corruption, while rura populations and those with higher education show views that are moderate. These results show the need for targeted anti-corruption campaigns to adress demographic differences and imporve governance accountability in Nigeria"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

library(modelsummary)
library(rstanarm)
library(dplyr)
library(here)
library(arrow)
library(bayesplot)
library(ggplot2)
library(arrow)
library(here)
library(patchwork)

@@

Understanding these demographic differences is important for designing effective anti-corruption campaigns. Making sure outreach efforts are tailored to the unique experiences and perceptions of different demographic groups can increase their effectiveness and increase public trust. For example, youth-focused campaigns that use social media platforms and urban outreach using investigative journalism can engage the people. At the same time, localized, culturally appropriate campaigns in rural areas can address the different needs of these populations. A the UNODC emphasizes, 'targeted and effective anti-corruption policies will improve likes of Nigerian citizens" [@afrobarometer2023].

The remainder of this paper is structured as follows; @sec-data detailed the data used for this paper, @sec-model introduces the Bayesian ordinal logistic regression model used to study the relationship between our variables of interest. @sec-results presents the findings from the model, detailing how different demographic factors like education and age affect corruption perception in Nigeria. @sec-discussion examines these results, discussing their implications for anti-corruption policies, weaknesses and next steps. @sec-survey-appendix explores the details of the Afrobarometer survey. @sec-model-details shows plots for Posterior predictive checks and Diagnostics as well as their analysis. @sec-survey-appendix details the survey methodology used in the Afrobaromenter data and @sec-ethics addresses very important ethical principles and the ethical challenges of surveying sensitive topics like corruption.

@@

Our primary measurement task is to capture public opinion towards
economic, political and socioeconomic issues and translate it into
actionable lintel. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses [@Afrobarometer2022codebook].

@@

Each survey question acts as a data proxy for these real-world
dynamics, reducing them to quantifiable variables while keeping
their conceptual integrity.This dataset shows a unique moment in Nigeria's socio-political
landscape, showing how citizens perceive democracy and governance in
the context of ongoing economic and political challenges. Alternative
datasets, such as Nigeria’s national statistics or World Bank survey
data, could provide economic indicators but do not have the granularity of
Afrobarometer’s public opinion data.

The data for this study was downloaded, cleaned, analyzed, modeled, and
visualized using R [@citeR], a statistical programming language.[@afrobarometer2023] served as the primary data source, providing importantsurvey data for analysis.***Telling Stories with Data*** [@tellingstories] was referenced for its code and methodologies. The
following packages were used:

@@

library(knitr)
library(kableExtra)

@@

The cleaned dataset used in this study is summarised in
@tbl-summary-statistics, which provides important statistics to provide an
overview of the dataset's structure and characteristics. The most
frequent age group among respondents is identified, representing the
predominant demographic in terms of age. The table also highlights the
urban and male percentages, which indicate the proportion of urban
residents (43.38%) and male respondents (49.68%) in the sample.

Further, the table breaks down the perceptions of corruption into four
categories—"None," "Some," "High," and "All"—with counts for each,
offering information into respondents' views on corruption in the
presidency. Additionally, the table details the distribution of
respondents across education levels, including categories such as
"Primary," "Secondary," "University," and "Postgraduate." Together,
these summary statistics provide a very detailed snapshot of the
dataset, enabling a clear understanding of its demographic and
categorical variables.

@@

The heatmap shows that respondents who completed secondary school or
some secondary schooling dominate the High Corruption and Some
Corruption categories. Respondents with higher education levels
(university and post-graduate) are distributed more evenly across the
corruption perception categories, with fewer concentrated in any
specific category. Conversely, respondents with no formal schooling or
only informal education are sparsely represented across all
categories.There is a notable concentration of corruption perceptions
among respondents with mid-level education (secondary schooling), while
those with higher education tend to have more varied perceptions. These
findings suggest a potential link between educational attainment and
experiences or attitudes toward corruption, particularly among those
with secondary-level education.

@@

The chart shows that rural and urban respondents share similar
patterns of corruption perception. The majority of respondents in both
groups fall under the Some Corruption and High Corruption categories,
with Some Corruption being slightly more prevalent. The All Corrupt
category is the third most common perception in both groups, while the
No Corruption category is consistently the least represented.

@@

@@

Higher education levels consistently show positive probabilities,
indicating that individuals with higher education levels are more
likely to perceive higher corruption levels. Living in rural areas has a
significantly negative probability, suggesting that rural residents are
less likely to perceive high corruption compared to urban residents.
Gender shows a moderate positive probability, with females slightly more
likely to perceive higher corruption levels than males. There seems to
be minimal variation observed across age groups, suggesting that age
does not substantially influence perceptions of corruption.

@@

Urban respondents are somewhat more likely than their rural counterparts
to perceive high corruption in the presidency. This is consistent with
findings that urban populations have greater access to investigative
reporting and public discourse on corruption, which can amplify
awareness and perceptions [@u4nigeria2023]. On the other hand, rural
respondents, who may rely on localized patronage networks for essential
services, often exhibit more neutral or unclear perceptions of
corruption, influenced by limited media access and dependency on these
informal systems. Campaigns should focus on leveraging urban platforms
like social media, community-based programs, and public dialogues to
engage important urban audiences, while also expanding rural outreach
through local radio, grassroots organizations, and educational
interventions tailored to these settings.

@@

This paper is focused on looking at trends in the corruption perception
between different demographics and seeing how these differences can be
used to maximize anti corruption campaigns and policy in
Nigeria.From this paper, we find different actionable strategies for
improving the effectiveness of anti-corruption campaigns in Nigeria.
First, targeting rural and urban populations differently might be
effective. Urban residents, who have greater access to investigative
media and public debates, are more likely to perceive high corruption.
Campaigns in urban areas should therefore focus on digital platforms
like social media and mobile apps, as well as partnerships with
investigative journalists and public accountability platforms to engage
this group effectively. In rural areas, where reliance on patronage
systems and limited media access shape perceptions, grassroots
communication strategies like community radio, local town halls, and
collaborations with trusted community leaders are better suited to
disseminate anti-corruption messaging.

Education also emerges as an important determinant of corruption perceptions.
Secondary school graduates exhibit the highest levels of awareness, this
suggests that civic education campaigns should focus on this demographic
to harness their increased engagement. Integrating anti-corruption
education into school curricula could early awareness and long-term
attitudinal changes.Youth engagement is very important, given that
younger age groups, particularly those aged 25–34, report the highest
levels of dissatisfaction with corruption. Social media campaigns
tailored to platforms popular with youth, together with interactive town
hall discussions and youth-led grassroots initiatives, could channel
this demographic's energy into organized anti-corruption efforts.

Lastly, rebuilding public trust in institutions should be a priority.
Transparent hiring practices, visible enforcement of anti-corruption
laws, and strengthened whistle blower protections are important steps to
showing institutional accountability. Localized outreach through
community leaders can also help fix the trust divide, particularly in
rural settings. Policymakers should put these findings into a
data-driven framework so they can adapt strategies based on demographic
trends and feedback.

@@

- Africa-Wide Generalization: While Afrobarometer provides national-level information, its findings cannot be generalized to the entire continent due to contextual variations.

@@

- **Inclusion of Marginalized Populations**: Future surveys could include anonymous digital surveys to take from institutionalized or transient populations.
- **Addressing Non-Response Bias**: Enhanced recruitment strategies, such as partnerships with community organizations, could increase participation among underrepresented groups.
- **Community Engagement**: Sharing results directly with communities through accessible formats (e.g., infographics or local town halls) would increase transparency and accountability.

@@

The ethical considerations underpinning Afrobarometer’s methodology ensure the credibility and validity of its findings while safeguarding respondent dignity. However, addressing remaining gaps—such as the exclusion of certain populations and challenges in conflict zones—would further enhance the ethical rigor of corruption perception research in Africa. By adhering to established ethical frameworks and remaining responsive to emerging challenges, studies like this can continue to give important information while maintaining the highest ethical standards.To do my part in ensuring ethical standards are upheld, i have included a datasheet for the Afrobarometer dataset.

@@

@fig-demo The figure provides an overview of demographic variables and corruption perceptions from the Afrobarometer Nigeria Round 9 survey. Rural residents constitute a larger share of the sample, aligning with Nigeria's predominantly rural demographic landscape. Corruption perception data shows that "High Corruption" and "Some Corruption" dominate public opinion. The education level distribution highlights the prevalence of basic education, with the majority of respondents having completed secondary school. Finally, the age distribution indicates that the 25–34 age group is the most represented cohort, suggesting that young adults play a crucial role in shaping perceptions of corruption and governance in Nigeria.
![Demographic Distributions of the Afrobarometer Data](combined_plot.png){#fig-demo width="120%"}
